Texture Mapping 3D Models of Indoor Environments with Noisy Camera Poses

### 2014-01-04 spie2014/spie2014.tex
down to 15, but no graphics
down to 14

\begin{figure}
  \centering
  \includegraphics[width=4in]{flowchart.jpg}
  \caption{The proposed texture mapping procedure\\}
  \label{fig:flowchart}
\end{figure}

@@

The backpack system used in this paper has a number of laser range
scanners as well as side-facing cameras taking photos at a rate of 5
Hz \cite{liu2010indoor}. In order to collect data, an operator wears
the backpack system and traverses the environment to be scanned,
walking in straight lines at a standard pace and making an effort to
walk parallel to all walls in the area. Multiple localization and
loop-closure algorithms are then applied to the raw data collected by
the onboard sensors \cite{chen2010indoor, kua2012loopclosure,
  liu2010indoor}, and the backpack is localized\footnote{In this
  paper, we use the terms localization and pose recovery
  interchangeably, in that they both refer to recovering position and
  orientation.}  over its data collection period. This involves
recovering the 6 degrees of freedom for the backpack itself as well as
the cameras rigidly mounted on it. These results are noisy, which
motivates the approaches in this paper. Once this is complete, the
data from the laser range scanners is used to generate a 3D point
cloud of the surrounding environment. This point cloud is used to
construct simple, coarse models \cite{sanchez2012point,
  turnerfloorplan}, as well as highly detailed models
\cite{turnerwatertight}.  Regardless of resolution, size, or detail,
the model, consisting of geometric surfaces in 3D space, along with
the set of images captured by the backpack's cameras and their noisy
3D poses, are the input to the texture mapping problem.

@@

Though some of these regions from high-resolution models are not
completely flat, a plane is fitted to each region for the purposes of
texturing. This plane corresponds to the largest contiguous section of
coplanar triangles, as most regions are characterized by small
outcroppings from a largely planar surface, such as objects on a table
or features protruding from a wall. The 3D surface composing a region
is then projected onto its calculated plane, and the resulting 2D
planar polygon is used as the texturing surface. This 2D planar
approximation provides a simple, yet approximately accurate surface
onto which images can be projected in order to generate
textures. However, 3D surfaces are still retained in order to achieve
more accurate results when performing occlusion checks during the
texturing process. To improve runtime during such occlusion checks,
the triangles within all 3D surfaces are also loaded into a k-d tree
for quick lookup. Thus, for high-resolution models, each region has
both 3D geometry stored in a k-d tree for occlusion purposes, and an
approximate 2D representation stored as the texturing surface.

At this point, we either have a series of planar regions from a low
resolution model, or a set of planar approximations for regions as
well as an associated k-d tree from a high resolution model. The task
now is to generate a texture for each region.

\section{Related Work}
\label{sec:relatedWork}
There are many existing approaches to stitching together multiple
images to produce a larger, seamless image \cite{szeliski2006image,
  agarwalapanoramas, wangmultipleviews, coorg1997matching,
  debevechybrid, bernardinimultiplescans}. Generally, parts of images
are matched to each other, either through direct comparisons at a
pixel level, or through feature detection and matching. Images are
then transformed to maximize matches, often by computing homographies
between pairs of images, or by iteratively adjusting camera poses in 1
to 6 degrees of freedom.

Feature matching has a number of advantages over direct matching that
make it more suitable for our often non-planar environments, and
rotational differences between camera poses
\cite{szeliski2006image}. Feature matching however, works best when
multiple unique visual references exist in the environment that can be
detected in multiple images. In contrast, indoor environments have a
high prevalence of bare surfaces, as well as repeating textures, such
as similar windows, doors, and wall-mounted decorations, that are
difficult to disambiguate. This lack of reliable reference points
often results in errors when matching images together.

Additionally, our datasets often contain long chains of images,
corresponding to long hallways and corridors as shown in Figure
\ref{fig:mosaic3D}, which leads to error accumulation when image
correspondences are not accurate. For example, when matching a long
chain of images through homography, a pixel in the $nth$ image must be
translated into the first image's coordinates by multiplying by the
$3\times3$ matrix $H_1 H_2 H_3 ... H_n$. Any error in one of these
homography matrices is propagated to all further images, resulting in
drift.

@@

In prior work, we have experimented with integrating image stitching
with the iterative global localization algorithm used to recover the 6
degrees of freedom of the backpack acquisition system detailed in
Section \ref{sec:dataAcquisition} \cite{liu2010indoor}. When run on
long chains of images however, especially where features are sparse,
this approach produces distorted textures, as seen in Figure
\ref{fig:mosaic3D}(a). Furthermore, this approach is not closed-form,
and its iterative camera adjustment process over large datasets leads
to prohibitively long computation time.

The AutoStitch software package, based on research in the related area
of panorama genoration, also performs homography-based alignment, with
additional provisions that attempt to reduce drift and increase
overall efficiency \cite{panorama2d, autostitch}. Though AutoStitch
performs well in small areas, it often induces drift in images,
such that global alignment to the environment geometry is
lost. Furthermore, its performance relies on the presence of a dense
set of features, and it has trouble aligning wall sections with even
short segments of bare texture. The example in Figure
\ref{fig:mosaic3D}(b) was generated after many rounds of manual
tuning; we have empirically found that areas with fewer visual
features or repeating texture patterns simply failed outright with
AutoStitch.

@@

\begin{figure}
  \centering \subfloat[][]{\includegraphics[width=3.4in,
    height=1.2in]{wall2_naive.jpg}}
  \subfloat[][]{\includegraphics[width=3.4in,
    height=1.2in]{wall2_naive_shift.jpg}}

  \centering \subfloat[][]{\includegraphics[width=3.4in,
    height=1.2in]{wall2_cache_shift.jpg}}
  \subfloat[][]{\includegraphics[width=3.4in,
    height=1.2in]{wall2_shortest_shift.pdf}}

  \centering \subfloat[][]{\includegraphics[width=3.4in,
    height=1.2in]{wall2_cache_shift_blend.pdf}}
  \subfloat[][]{\includegraphics[width=3.4in,
    height=1.2in]{wall2_shortest_shift_blend.pdf}}

  \centering \subfloat[][]{\includegraphics[width=3.4in,
    height=3.4in]{wall2_cache_comparison.png}}
  \subfloat[][]{\includegraphics[width=3.4in,
    height=3.4in]{wall2_shortest_comparison.png}}

  \caption{(a) Tile-based texturing; (b) Tile-based texturing after
    image alignment; (c) Tile-based texturing after image alignment
    with caching; (d) Shortest path texturing after image alignment;
    (e,f) Blending applied to (c) and (d); (g,h) Zoomed in views of
    discontinuities in (e) vs. in (f).}
  \label{fig:compareAll}
\end{figure}

@@

\begin{figure}
  \centering
  \subfloat[][]{\includegraphics[width=3.3in]{exposureDiff1.png}}
  \subfloat[][]{\includegraphics[width=3.3in]{exposureDiff2.png}}
  \caption{(a) A ceiling texture composed of images taken with varying
    exposures, with significant brightness differences. (b) Exposure
    compensation applied to the same set of images from (a) by
    applying computed gains to each image}
  \label{fig:exposureDiff}
\end{figure}

@@

After linear alpha blending across overlapping regions, the texture
mapping process is complete. Figures \ref{fig:compareAll}(e) and
\ref{fig:compareAll}(f) show the blended versions of Figures
\ref{fig:compareAll}(c) and \ref{fig:compareAll}(d) respectively. The
remaining images in Figure \ref{fig:compareAll} highlight differences
between the two methods, showing that Figure \ref{fig:compareAll}(f)
has the best visual quality and the best texturing approach among the
textures in Figure \ref{fig:compareAll}.

>>>>

\begin{figure}
  \centering
  \includegraphics[width=6.2in]{flowchart2.jpg}
  \caption{The proposed texture mapping procedure\\}
  \label{fig:flowchart}
\end{figure}

@@

The backpack system used in this paper has a number of laser range
scanners as well as side-facing cameras taking photos at a rate of 5
Hz \cite{liu2010indoor}. In order to collect data, an operator wears
the backpack system and traverses the environment to be scanned,
walking in straight lines at a standard pace and making an effort to
walk parallel to all walls in the area. Multiple localization and
loop-closure algorithms are then applied to the raw data collected by
the onboard sensors \cite{chen2010indoor, kua2012loopclosure,
  liu2010indoor}, and the backpack is localized\footnote{In this
  paper, we use the terms localization and pose recovery
  interchangeably, in that they both refer to recovering position and
  orientation.}  over its data collection period. This involves
recovering the 6 degrees of freedom for the backpack itself as well as
the cameras rigidly mounted on it. These results are noisy, which
motivates the approaches in this paper. Once this is complete, the
data from the laser range scanners is used to generate a 3D point
cloud of the surrounding environment. This point cloud is used to
construct simple, coarse models \cite{sanchez2012point,
  turnerfloorplan}, as well as highly detailed models
\cite{turnerwatertight}.  Regardless of resolution or size,
the model, consisting of geometric surfaces in 3D space, along with
the set of images captured by the backpack's cameras and their noisy
3D poses, are the input to the texture mapping problem.

@@

For the purposes of texturing, it is convenient to treat each region
as planar. To accomplish this, a plane is fitted to each region. This
plane corresponds to the largest contiguous section of coplanar
triangles, as most regions are characterized by small outcroppings
from a largely planar surface, such as objects on a table or features
protruding from a wall. The 3D surface composing a region is then
projected onto its calculated plane, and the resulting 2D planar
polygon is used as the texturing surface. This 2D planar approximation
provides a simple, yet approximately accurate surface onto which
images can be projected in order to generate textures. However, 3D
surfaces are still retained in order to achieve accurate results when
performing occlusion checks, and of course are used to generate the
final textured model. Each region now has a set of triangles
representing the original model geometry, as well as a planar
approximation. The task now is to generate a texture for each region.

\section{Related Work}
\label{sec:relatedWork}

@@

There are many existing approaches to stitching together multiple
images to produce a larger, seamless image \cite{szeliski2006image,
  agarwalapanoramas, wangmultipleviews, coorg1997matching,
  debevechybrid, bernardinimultiplescans}. Generally, parts of images
are matched to each other, by detecting feature points and identifying
matches. Images are then transformed to maximizally align matches,
often by computing homographies between pairs of images, or by
iteratively adjusting camera poses in 1 to 6 degrees of freedom.

Feature matching works best when unique visual references exist in the
environment that can be detected in multiple images. Unfortunately,
many indoor environments have a high prevalence of bare surfaces as
well as repeating textures, such as with windows and
doors. Additionally, our datasets often contain long 1-dimensional
chains of images, such as in Figure \ref{fig:mosaic3D}, which often
leads to error accumulation. For example, when matching a long chain
of images via homography matrices, a pixel in the $nth$ image must be
translated into the first image's coordinates by multiplying by the
$3\times3$ matrix $H_1 H_2 H_3 ... H_n$. Errors in each of these
homography matrices is propagated to all further images, resulting in
drift.

Two methods based on feature matching and homography calculation are
shown in Figure \ref{fig:mosaic3D}. Figure \ref{fig:mosaic3D}(a)
depicts an attempt at integrating image stitching with the iterative
global localization algorithm used to localize the backpack system
detailed in Section \ref{sec:dataAcquisition}
\cite{liu2010indoor}. Figure \ref{fig:mosaic3D}(b) depicts an output
image from AutoStitch, which is software based on research in the
related area of panorama generation \cite{panorama2d,
  autostitch}. Both methods were thoroughly tuned for the shown
example, but due to the factors mentioned in the previous paragraph,
they still show significant amounts of drift and distortion, as well
as loss of alignment to the environment geometry.

@@

\begin{figure}
  \centering \subfloat[][]{\includegraphics[width=0.33\textwidth,
    height=0.8in]{wall2_naive.jpg}}
  \subfloat[][]{\includegraphics[width=0.33\textwidth,
    height=0.8in]{wall2_naive_shift.jpg}}
  \subfloat[][]{\includegraphics[trim=1.2cm 0cm 0cm 0cm,
        clip=true,width=0.33\textwidth,
    height=0.8in]{wall2_cache_shift.jpg}}

  \centering \subfloat[][]{\includegraphics[width=0.33\textwidth,
    height=0.8in]{wall2_shortest_shift.pdf}}
  \subfloat[][]{\includegraphics[width=0.33\textwidth,
    height=0.8in]{wall2_cache_shift_blend.pdf}}
  \subfloat[][]{\includegraphics[width=0.33\textwidth,
    height=0.8in]{wall2_shortest_shift_blend.pdf}}

  \caption{(a) Tile-based texturing; (b) Tile-based texturing after
    image alignment; (c) Tile-based texturing after image alignment
    with caching; (d) Shortest path texturing after image alignment;
    (e,f) Blending applied to (c) and (d).}
  \label{fig:compareAll}
\end{figure}

@@

\begin{figure}
  \centering
  \subfloat[][]{\includegraphics[width=2.5in, height=0.9in]{exposureDiff1.png}}
  \hspace{0.5in}
  \subfloat[][]{\includegraphics[width=2.5in, height=0.9in]{exposureDiff2.png}}
  \caption{(a) A ceiling texture composed of images taken with varying
    exposures, with significant brightness differences. (b) Exposure
    compensation applied to the same set of images from (a) by
    applying computed gains to each image}
  \label{fig:exposureDiff}
\end{figure}

@@

After linear alpha blending across overlapping regions, the texture
mapping process is complete. Figures \ref{fig:compareAll}(e) and
\ref{fig:compareAll}(f) show the blended versions of Figures
\ref{fig:compareAll}(c) and \ref{fig:compareAll}(d) respectively. By
comparing all images, it is clear that Figure \ref{fig:compareAll}(f)
has the best visual quality and the best texturing approach among the
textures in Figure \ref{fig:compareAll}.
